---
title: 'Tidy Tuesday Week 37: NLP and Friends'
author: mcnakhaee
date: '2020-10-15'
slug: []
categories:
  - NLP
  - R
  - social network analysis
tags:
  - R
  - NLP
subtitle: ''
summary: ''
authors: []
lastmod: '2020-10-15T10:28:36+02:00'
featured: no
draft: true
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

I experience a few difficult and sad moments in my life in the past few months and it was not particularly an easy job for me to cope with these hard moments. However, I found a medicine to fight the sadness that was caused by these painful experience: watching Friends! I have watched Friends many many times but still it can bring smile to my face. Every time I feel being sad, I just start to watch several episodes on Friends, one after another and the laughter does the rest!

So, when I realized that the dataset we are going to work with in the 37th week of 2020 Tidy Tuesdays project, I was absolutely thrilled and excited to get my hands on this dataset and analyze it. As a big fan of Friends, many ideas popped  into my head on how I can make an interesting analysis on this dataset. In this post, I will go through three of them:
1. In Which locations (scenes) Friends was most filmed?
2. How characters in the show interacted with each other?
3. What kinds of actions female and male characters in the show were doing?

I will addressed these three questions in the rest of this post!




```{r message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	eval = FALSE
)
myFont <- 'Rubik'
windowsFonts(Rubik  = myFont)
extrafont::loadfonts(device="win")
myFont <- "Poppins Light"
windowsFonts(`Poppins Light`  = myFont)
library(reticulate)
use_condaenv('spacy_condaenv')
#conda_list()
Sys.setenv(RETICULATE_PYTHON = 'C:\\Users\\mcnak\\AppData\\Local\\Continuum\\anaconda3\\envs\\spacy_condaenv\\python.exe')

```


```{r}

library(tidyverse)
library(ggthemes)
library(tidytuesdayR)
library(rsyntax)
library(spacyr)
library(tidytext)
library(tidygraph)
library(igraph)
library(ggraph)
library(hrbrthemes)
library(DT)
spacy_initialize(model = "en_core_web_lg")

theme_set(  theme_fivethirtyeight()) 
theme_update(
    text = element_text(family = 'Poppins Light', color = '#774936'),
    plot.title = element_text(margin = margin(t = 10, b = 10)),
    plot.subtitle = element_text(margin = margin(b = 20)),
    axis.text.y = element_blank(),
    plot.background = element_rect(fill = '#FFFAF0'),
    panel.background = element_rect(fill = '#FFFAF0'),
    legend.position = 'none',
    panel.grid.major  = element_line(
      color = '#eddcd2',
      size = 0.5,
      linetype = 'dashed'
    ),
    panel.grid.minor = element_line(
      color = '#eddcd2',
      size = 0.01,
      linetype = 'dashed'
    )
  )
```

Let us load the Friends dataset using the handy `tt_load` function from the `tidytuesdayR` package!

```{r}
tt <- tt_load("2020-09-08")
friends_df <- tt$friends 
```

It returns three tables but since the data that I need , I will only be working with the friends transcript dataset.
```{r}
friends_df %>% 
  head(20) %>% 
  datatable()
  
```



```{r}
friends_df <- tt$friends 
```

### Top Locations 

Now let's find out which locations this show was filmed the most. This information is available in the `text` column.
In this column,Usually the location where the scene took place comes after the `Scene:` string . So, we only need to find and filter rows that have this string. Unfortunately, this information is not available for the first and the last two seasons of Friends!

```{r}
friends_df %>%
  filter(str_detect(text, fixed('Scene:'))) %>% 
  head() %>% 
  datatable()
```




```{r }
scences <- friends_df %>%
  filter(str_detect(text, fixed('Scene:'))) %>%
  rowwise() %>%
  mutate(
    location = if_else(season <9,str_split(text, ',')[[1]][1],str_split(text, fixed('.'))[[1]][1]),
    location = str_remove_all(location, 'erm'),
    location = str_remove_all(location, fixed('erm')),
    location = str_remove(location, fixed('[Scene:')) ,
    location = str_trim(location)
  ) %>% 
  filter(!str_detect(location, fixed(']')),
         !location %in% c('Monica', 'Chandler'))

scenes_location_count <- scences %>%
  count(location, sort = TRUE) %>%
  filter(n > 10 )


```
Now let's sort and look at the top locations in Friends:
```{r}
    
scenes_location_count  %>% 
  DT::datatable()
```
So, Central Perk was the location where most scenes took place!

```{r fig.height=5,fig.width=10}
 scenes_location_count %>% 
  ggplot(aes(x = fct_reorder(location,n),n)) +
  geom_col(fill = '#ef476f') +
  coord_flip() +
  labs(title = 'Most Filmed Locations in Friends' ) +
  theme_fivethirtyeight() +
  theme(legend.position = 'none', 
        axis.text.x = element_text(angle = 90))
```
But let's look at how many times each different locations

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
scences %>% 
  count(season,location,sort = TRUE) %>% 
  filter(n>5) %>% 
  ggplot(aes(x= season,y=n)) +
  geom_line(color = 'red',size = 1)+
  geom_point(color = 'red',size = 3)+
  labs(title = 'Where Was Freinds Filmed?',subtitle = '') +
  facet_wrap(~location) 
  
  
```







it inspired me to perform the same analysis on the Friends' transcript. However, my approach is slightly different from what Julia Sigles employed. She used bigrams to find a combination of pronouns and verbs while My approach uses named entity recognition (NER) form the natural language processing literature. In my previous posts, I used Spacy and Python to extract named entities from text data, but in this post I will exclusively use 
the `spacyr` package in R (which behind the scene uses Spacy and python).


```{r}
female_characters <- c('monica','rachel','phoebe','emily','carol','janice','susan','kathy','elizabeth','she')
male_characters <- c('ross','joey','chandler','mike','richard','gunther','paolo','bob','paul','frank','he')


male_speakers <- c('Ross Geller','Chandler Bing','Joey Tribbiani','Richard Burke','Mike Hannigan','Frank Buffay Jr.','Jack Geller','David','Peter Becker','Eddie Menuek','Gunther','Paul Stevens')

female_speakers <- c('Rachel Green','Phoebe Buffay','Janice Litman Goralnik','Monica Geller','Emily Waltham','Charlie Wheeler','Judy Geller','Amy Green','Carol Willick','Mona')
```

I start by separating the text column into sentences and give each sentence a unique id which I use later for joining dataframes.

```{r}
friends_sentences <- friends_df %>%
  unnest_tokens(sentence, text, token = 'sentences') %>%
  rownames_to_column() %>%
  rename(doc_id = rowname,
         text = sentence) 
friends_sentences %>%
  tail() %>% 
  datatable()
```

The main function that I am going to use from `spacyr` is  the 'spacy_parse' function which can return named entities, dependencies, part-of-speech tags, etc. By default, it does not give us dependencies so we need to set `dependency = TRUE`.

```{r}
tq <- tquery(pos = "VERB", label = "verb", children(relation = "nsubj", label = "subject"))
tokens_df_dep <- friends_sentences %>% 
  spacy_parse(dependency = TRUE) %>%
  annotate_tqueries("annotation", query_name = tq) %>%
  filter(!is.na(annotation))


```

Now I join tokens that I created using spacy and the firends sentences dataframe.

```{r eval=FALSE, include=FALSE}
friends_entities <- friends_sentences %>% 
  inner_join(tokens_df_dep,by = 'doc_id') %>% 
  filter(!token %in% c('Bye','kiddin','God','Sting','Bijan')) %>% 
  mutate(token = case_when(str_detect(str_to_lower(token),'rach') ~ 'Rachel',
                           str_detect(str_to_lower(token),'joe')~ 'Joey',
                            str_detect(str_to_lower(token),'pheebs')  ~ 'Phoebe',
                           TRUE ~ token))

#!speaker %in% c('Scene Directions'),
#str_detect(entity,'PERSON_B'),
```


I will pull the documnets id that:

token

```{r eval=FALSE, include=FALSE}


female_ids <- friends_entities %>% 
  filter((token %in% female_characters | ( speaker %in% female_speakers & token =='i')), annotation =='subject' ) %>% 
  pull(doc_id)
  
male_ids <- friends_entities %>% 
  filter((token %in% male_characters| (speaker %in% male_speakers & token =='i')), annotation =='subject'  ) %>% 
  pull(doc_id)
female_verbs <- friends_entities %>% 
  filter(doc_id %in% female_ids,pos == 'VERB',relation =='ROOT'  ) %>% 
  unnest_tokens(word,lemma) %>% 
  anti_join(stop_words) %>% 
  count(word,sort = TRUE) %>% 
  mutate(pronoun = 'she')
male_verbs <- friends_entities %>% 
  filter(doc_id %in% male_ids,pos == 'VERB',relation =='ROOT') %>% 
  unnest_tokens(word,lemma) %>% 
  anti_join(stop_words) %>% 
  count(word,sort = TRUE) %>% 
  mutate(pronoun = 'he')
```

```{r eval=FALSE, include=FALSE}
verb_log <- female_verbs %>% 
  bind_rows(male_verbs) %>%
  filter(word!='singe') %>% 
  pivot_wider(names_from = pronoun,values_from = n,values_fill = list(n = 0)) %>% 
  mutate(total = he + she,
         he = (he+1)/sum(he + 1),
         she = (she + 1)/sum(she + 1),
         logratio = log2(she / he)) %>% 
  filter(total > 15) %>% 
  group_by(logratio < 0) %>%
  top_n(15,abs(logratio)) %>% 
  ungroup() %>% 
  mutate(word = reorder(word,logratio),
         nchars = nchar(as.character(word))) 
```




```{r eval=FALSE, fig.height=10, fig.width=10, include=FALSE}
female_arrows <- tibble(
  x1 = c(23),
  x2 = c(23),
  y1 = c(1),
  y2 = c(1.8)
)
male_arrows <-  tibble(
  x1 = c(9),
  x2 = c(9),
  y1 = c(-1),
  y2 = c(-1.8)
)
female_color = '#ef476f'
male_color = '#118ab2'
g <- ggplot() +
  geom_col(data = verb_log , aes(word, logratio, fill =  logratio < 0)) +
  geom_text(
    data = verb_log ,
    aes(
      label = word,
      x = word,
      y = if_else(logratio < 0, logratio + 0.03*nchars, logratio  - 0.03*nchars)
    ),
    color = 'white',
    family = 'Rubik',
    face = 'bold',
    #label.padding = 0.05
  ) +
  
  coord_flip() +
  scale_y_continuous(limits = c(-2, 2)) +
  scale_fill_manual(values = c(female_color, male_color)) +
  ggplot2::annotate(
    "text",
    x = 24,
    y = 1.5,
    label = 'More Female',
    family = 'Rubik',
    color = female_color,
    size = 5,
    vjust = 0.5,
    hjust = 0.6
  ) +
  ggplot2::annotate(
    "text",
    x = 10,
    y = -1.34,
    label = 'More Male',
    family = 'Rubik',
    color = male_color,
    size = 5,
    vjust = 0.5,
    hjust = 0.6
  ) +
  
  geom_curve(
    data = female_arrows,
    aes(
      x = x1,
      y = y1,
      xend = x2,
      yend = y2
    ),
    arrow = arrow(length = unit(0.07, "inch")),
    size = 2,
    color = female_color,
    curvature = 0
  ) +
  
  geom_curve(
    data = male_arrows,
    aes(
      x = x1,
      y = y1,
      xend = x2,
      yend = y2
    ),
    arrow = arrow(length = unit(0.07, "inch")),
    size = 2,
    color = male_color,
    curvature = 0
  ) +
  labs(
    title = 'She Lies, He Listens',
    subtitle = 'This plot shows what common verbs were used to describe male and female characters in Friends',
    fill = '',
    caption =  'visualization: mcnakhaee.com'
  ) 

g
```



### Friends' Network


```{r eval=FALSE, include=FALSE}
tokens <- tt$friends %>%
  rownames_to_column() %>%
  rename(doc_id = rowname) %>%
spacy_parse(pos = FALSE,tag = FALSE)
```



```{r eval=FALSE, include=FALSE}
tokens_speakers_count <- friends_entities %>% 
  count(speaker,token,sort = TRUE) %>% 
  filter(n >5) %>% 

speakers <- tokens_speakers_count %>% 
  distinct(speaker) %>% 
  pull()
```



```{r eval=FALSE, fig.height=15, fig.width=15, include=FALSE}
traph_df <- tokens_speakers_count %>% 
  rename(from = speaker,
         to = token) %>% 
  as_tbl_graph(directed = TRUE) %>%
  activate(nodes) %>%
  mutate( deg_cent = log10(centrality_degree() +1) + 1,
         is_speaker = if_else(name %in% speakers,TRUE,FALSE) )

traph_df %>%
  ggraph(layout = 'nicely') +
  geom_edge_link(aes(width = n,alpha = n),color = '#f9dcc4') +
  geom_node_text(aes(label = name,color = is_speaker),repel = TRUE,family = 'Rubik') +
  geom_node_point(aes(size = deg_cent,color = is_speaker),alpha = 0.5) +
  scale_color_tableau() +
  theme_graph() + 
    theme(
    plot.title = element_text(
      family = 'American Typewriter',
      face = "bold",
      size = 60,
      margin = ggplot2::margin(0, 0, 20, 0),
      hjust = 0
    ),
    legend.position = 'none',
    plot.subtitle = element_text(
      family = 'American Typewriter',
      size = 40,
      margin = margin(0, 0, 60, 0),
      hjust = 0
    ),
    strip.text = element_text(
      family = 'American Typewriter',
      size = 40,
      #margin = margin(0, 0, 60, 0),
      hjust = 0.5
    ),
    
        plot.caption = element_text(
      family = 'American Typewriter',
      size = 30,
      #margin = margin(0, 0, 60, 0),
      #hjust = 0.5
    ),
    strip.background = element_rect(color = 'gray50',
                                    fill = "#fc9d9d"),
  ) 

```



```{r eval=FALSE, fig.height=40, fig.width=20, include=FALSE}
tokens_speakers_season_count <- friends_entities %>% 
  count(season,speaker,token,sort = TRUE) %>% 
  filter(n >15)
  
  
graph_df <- tokens_speakers_season_count %>% 
  rename(from = speaker,
         to = token) %>% 
  as_tbl_graph(directed = TRUE) %>%
  activate(nodes) %>%
  mutate(deg_cent= centrality_degree(mode = 'out') + centrality_degree(mode = 'in') ,
         is_speaker = if_else(name %in% speakers,TRUE,FALSE) )  



graph_df %>%
  ggraph(layout = 'nicely') +
  geom_edge_link(aes(width = n,alpha = n),color = '#f9dcc4') +
  geom_node_text(aes(label = name,color = is_speaker),repel = TRUE,size = 4,family = 'Rubik') +
  geom_node_point(aes(size = deg_cent+2,color = is_speaker),alpha = 0.5) +
  scale_color_tableau() +
  facet_edges(~season,nrow = 5) +
  theme_graph() + 
    theme(
    plot.title = element_text(
      family = 'American Typewriter',
      face = "bold",
      size = 60,
      margin = ggplot2::margin(0, 0, 20, 0),
      hjust = 0
    ),
    legend.position = 'none',
    plot.subtitle = element_text(
      family = 'American Typewriter',
      size = 40,
      margin = margin(0, 0, 60, 0),
      hjust = 0
    ),
    strip.text = element_text(
      family = 'American Typewriter',
      size = 40,
      #margin = margin(0, 0, 60, 0),
      hjust = 0.5
    ),
    
        plot.caption = element_text(
      family = 'American Typewriter',
      size = 30,
      #margin = margin(0, 0, 60, 0),
      #hjust = 0.5
    ),
    strip.background = element_rect(color = 'gray50',
                                    fill = "#fc9d9d"),
  ) 
```

